---
title: "pubh7462_hw1_chri3744"
author: "Taylor Christian"
date: "1/29/2022"
output: github_document
---

```{r setup, include=FALSE}
#Load all the good stuff
library(tidyverse)

#Working directory for .RMD
knitr::opts_knit$set(echo = TRUE,
                     root.dir = rprojroot::find_rstudio_root_file())

#Controlling figure output in markdown
knitr::opts_chunk$set(
#  fig.height =   
  fig.width = 6,
#  fig.asp = .5,
  out.width = "90%",
#  out.height = 
 fig.align = "center",
  cache = FALSE,
  echo  = TRUE
)

#Set Theme for ggplot2 - centers title and legend at bottom by default
theme_set(theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom"))

#Set Scientific notation output and decimal places for knitr
options(scipen = 999)
options(digits = 4)
```


# Problem 2.1
```{r include = TRUE}
#Creating tibble
hw1_tb <- tibble(
  x = rnorm(1000, mean = 0, sd = sqrt(1)),
  y = rnorm(1000, mean = 1, sd = sqrt(2)),
  sum_indicator = ifelse(x + y > 5, "Yes", "No")
)

#Converting sum_indicator to factor
hw1_tb <- mutate(hw1_tb, sum_indicator = as.factor(sum_indicator))

#Reordering the factor so Yes is before No
hw1_tb <- mutate(hw1_tb, sum_indicator = forcats::fct_relevel(sum_indicator, "Yes"))

#Creating ggplot
hw1_tb.gg <- hw1_tb %>%
  ggplot(aes(x = x, y = y, color = sum_indicator)) +
  geom_point()

#Displaying ggplot
hw1_tb.gg +
  #Labeling ggplot
  labs(x = "X ~ N(0,1)", y = "Y ~ N(1,2)", title = "Plotting Normal Random Samples Against Eachother") +
  #Adding legend title
  scale_colour_viridis_d("X + Y > 0.5")
```


# Problem 2.2
```{r include = TRUE}
penguin.df <- read_rds("./data/penguin.RDS")
```

## Problem 2.2.1

### Case definition
Each row represents an observation of penguin species, Palmer Archipelago island, size measurements(flipper length, body mass, and bill dimensions), and sex.\

### Number of Observations
There were `r nrow(penguin.df)` observations made for this data.\

### Number of Variables/Features
There are `r ncol(penguin.df)` variables for each observation.\

### Meaning of Each Variable
**species**: a factor denoting penguin species (Adelie, Chinstrap, and Gentoo).\
**island**: a factor denoting island in Palmer Archipelago, Antartica (Biscoc, Dream, or Torgersen).\
**bill_length_mm**: a number denoting bill length (millimeters).\
**bill_depth_mm**:  a number denoting bill depth (millimeters).\
**flipper_length_mm**: an integer denoting flipper length (millimeters).\
**body_mass_g**: an integer denoting body mass (grams).\
**sex**: a factor denoting penguin sex (female, male).\
**year**: an integer denoting the study year (2007, 2008, or 2009).\

### Variable Means

#### Flipper length mean and SD
The observed mean flipper length is `r round(mean(penguin.df$flipper_length_mm, na.rm = TRUE), digits = 2)` and its SD is `r round(sd(penguin.df$flipper_length_mm, na.rm = TRUE), digits = 2)`.

#### Bill length mean and SD
The observed mean flipper length is `r round(mean(penguin.df$bill_length_mm, na.rm = TRUE,), digits = 2)` and its SD is `r round(sd(penguin.df$bill_length_mm, na.rm = TRUE), digits = 2)`.

## Problem 2.2.2

```{r include = TRUE, message = FALSE, warning = FALSE}
#Creating ggplot
penguin.gg <- penguin.df %>%
  ggplot(aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point()

#Displaying ggplot
penguin.gg +
  #Labeling ggplot
  labs(x = "Bill Length (mm)", y = "Flipper Length (mm)", title = "Penguin Bill Length vs Flipper Length") +
  #Adding legend title
  scale_colour_viridis_d("Species")
```

```{r include = TRUE, message = FALSE, warning = FALSE}
#Creating ggplot
penguin.gg <- penguin.df %>%
  ggplot(aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point()

#Displaying ggplot
penguin.gg +
  #Labeling ggplot
  labs(x = "Bill Length (mm)", y = "Flipper Length (mm)", title = "Penguin Bill Length vs Flipper Length by Species") +
  #Adding legend title
  scale_colour_viridis_d("Species") +
  #Splitting by species
  facet_grid(vars(species))
```

### Data Trends
From looking at the above plots, it can be seen that average bill length is shorter in the Adelie penguins than in the other two species observed.  Additionally, average flipper length appears higher in the observed Gentoo penguins than the other two species observed.  In all three species there appears to be a positive correlation between flipper length and bill length.